---
output: github_document
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  dev = "png",
  dpi = 200,
  fig.align = "center",
  knitr::opts_chunk$set(comment = NA)
  )
library(qgraph)
library(bootnet)
library(CliquePercolation)
```

[![CRAN Version](http://www.r-pkg.org/badges/version/CliquePercolation)](https://cran.r-project.org/package=CliquePercolation)
[![Downloads](https://cranlogs.r-pkg.org/badges/CliquePercolation)](https://cran.r-project.org/package=CliquePercolation)
[![Travis build status](https://travis-ci.com/LangeJens/CliquePercolation.svg?branch=master)](https://travis-ci.com/LangeJens/CliquePercolation)

# CliquePercolation

**CliquePercolation** is an **R** package that entails multiple functions related to the clique percolation community detection algorithms for undirected, unweighted networks and for undirected, weighted networks. As various sciences analyze structural properties of networks, the **CliquePercolation** package can be useful in many fields such as physics, biology, psychology, computer science, or sociology.

## Installation

To use **CliquePercolation**, you first need to install **R** (find download page [here](https://cran.r-project.org/)). Moreover, I recommend to use **R** via *RStudio*, an environment that facilitates the use of **R** in multiple ways (find download page [here](https://rstudio.com/)).

The developmental version of **CliquePercolation** is available on GitHub. You can download it from GitHub directly with the help of the **devtools** package in **R**:


```{r, eval = FALSE}
install.packages("devtools")
devtools::install_github("LangeJens/CliquePercolation", build_vignettes = TRUE)
```


You can download the current stable version from CRAN:


```{r, eval = FALSE}
install.packages("CliquePercolation")
```

## Community Guidelines

If you have recommendations for improving **CliquePercolation**, want to report issues or problems with the package, or need support, you can either open an [issue](https://github.com/LangeJens/CliquePercolation/issues) on GitHub or contact me directly via e-mail (<lange.jens@outlook.com>). If you consider contributing to **CliquePercolation**, you can also submit a [pull request](https://github.com/LangeJens/CliquePercolation/pulls).

## Illustrative Examples

The following includes two examples -- one for an undirected, weighted network and one for an undirected, unweighted network -- to illustrate parts of the package with real data. For a more extensive description of (a) the clique percolation algorithms for undirected, weighted as well as undirected, unweighted networks and (b) the workflow of the package with multiple, fictitious examples, see the package vignette, e.g., by running `vignette("CliquePercolation")` after installing and loading (`library(CliquePercolation)`) the package. Furthermore, the documentation of each function covers various other standalone examples with fictitious and real data that can be used to verify their functionality. An overview of all functions and links to their documentations are available by running `?'CliquePercolation'`.

### The Obama network

The package includes a data set with 10 evaluative reactions toward Barack Obama from 5,914 participants (for details see `?Obama`). This data set can be used to estimate a Gaussian Graphical Model (i.e., a regularized partial correlation network), which is an undirected, weighted network. Subsequently, the **CliquePercolation** package allows analyzing the community structure of the network.

First, we will load the data, estimate the network with the package **bootnet**, and plot the network with the package **qgraph**.

```{r, echo = TRUE, warning = FALSE, message = FALSE}
#load data
data(Obama)

#estimate network
net <- bootnet::estimateNetwork(Obama, default = "EBICglasso", missing = "pairwise")

#plot network
graph <- plot(net, layout = "spring")
```

Second, we optimize *k* and *I* values for running the clique percolation algorithm. We do so by determining the entropy values of community partitions for a range of *k* and *I* values and then applying a permutation test to determine which of these combinations produce community partitions that are more surprising than already expected by chance.

```{r, echo = TRUE, message = FALSE}
#determine entropy for a range of k and I values
threshold <- cpThreshold(graph, method = "weighted",
                         k.range = 3:4,
                         I.range = seq(0.1, 0.5, 0.01),
                         threshold = "entropy")

#use permutation test to determine which k and I are optimal
permute <- cpPermuteEntropy(graph, cpThreshold.object = threshold,
                            ncores = 2, seed = 4186)
permute
```

Third, we run the clique percolation algorithm with the optimized values. That is, *k = 3* and *I = 0.14* produce the highest entropy with the smallest number of isolated nodes that is more surprising than already expected by chance.

```{r, echo = TRUE}
#run clique percolation algorithm with optimal k and I
cpk3I.14 <- cpAlgorithm(graph, k = 3, I = 0.14, method = "weighted")

#print results overview
cpk3I.14

#show detailed summary of results
summary(cpk3I.14)
```

Finally, we plot the community partition with three communities and three shared nodes onto the original network.

```{r, echo = TRUE}
#plot community partition on original network plot
coloredGraph <- cpColoredGraph(graph,
                               list.of.communities = cpk3I.14$list.of.communities.labels,
                               layout = "spring", theme = "colorblind")
```

The plot indicates that feelings and beliefs partly split up in separate, yet overlapping communities.

### The immunoglobulin network

The package also includes an undirected, unweighted network of interactions of 1,316 amino-acids (for details see `?immuno`), which we can also analyze with the **CliquePercolation** package.

First, we will load the data and plot the network with the package **qgraph**.

```{r, echo = TRUE, warning = FALSE, message = FALSE}
#load data
data(immuno)

#plot network
graph <- qgraph::qgraph(immuno, layout = "spring", vsize = 1)

```

Second, we optimize *k* for running the clique percolation algorithm. We do so by determining the ratio of the largest to the second largest community and the chi values for a range of *k* values to determine whether a community partition leads to a ratio larger than 2 (and if yes, which is the first community partition to exceed this threshold) and/or to the highest chi value with a low number of isolated nodes.

```{r, echo = TRUE, message = FALSE}
#determine k for clique percolation
threshold <- cpThreshold(graph, method = "unweighted",
                         k.range = 4:5,
                         threshold = c("largest.components.ratio", "chi"))
threshold
```

Third, we run the clique percolation algorithm with the optimized value. That is, *k = 4* produces the highest chi with a low number of isolated nodes.

```{r, echo = TRUE}
#run clique percolation algorithm with optimal k
cpk4 <- cpAlgorithm(graph, k = 4, method = "unweighted")

#print results overview
cpk4
```

Finally, we plot the community graph, namely a network in which each node represents a community and edges indicate the number of nodes the communities share. Larger nodes represent larger communities.

```{r, echo = TRUE}
#plot community graph
communityGraph <- cpCommunityGraph(cpk4$list.of.communities.numbers,
                                   node.size.method = "proportional",
                                   max.node.size = 10,
                                   theme = "colorblind",
                                   layout = "spring",
                                   repulsion = 0.98)
```

The plot indicates that the interactions of amino-acids in immunoglobulin are structurally organized in three blocks of interrelated communities. All three blocks are dominated by a few large communities.